What is long-distance dispersal? And a taxonomy of dispersal events\\ \vspace{2cm}

### 2016-10-10 Manuscript/MS- What is LDD.tex
Edits rev #2.

1. Dispersal is a key individual-based process influencing many life-history attributes, scaling up to population-level properties (e.g., metapopulation connectivity). A persistent challenge in dispersal ecology has been the robust characterization of dispersal functions (kernels), a fundamental tool to predict how dispersal processes respond under global change scenarios. Especially the rightmost tail of these functions, i.e. the long-distance dispersal (LDD) events, are difficult to characterize empirically and to model in realistic ways. \\
2. But, when is it a LDD event? In the specific case of plants, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event. Here we discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration. \\
3. Strict-sense long distance dispersal involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals. Combinations of propagule movements within/outside these two spatial reference frames results in four distinct modes of LDD. \\
4. \textit{Synthesis.} We expect truncation of seed dispersal kernels to have multiple consequences on demography and genetics, following to the loss of key dispersal services in natural populations. Irrespective of neighborhood sizes, loss of LDD events may result in more structured and less cohesive genetic pools, with increased isolation-by-distance extending over broader areas. Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.\\

@@

We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of their frequency and extent could be comparable across different study systems. We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses. These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD. We start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds. We then use a case study to assess differential contributions of animal frugivores performing LDD.

Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events. We propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives. We also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds. Finally, we highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.

@@

Here we propose an explicit definition of LDD and what constitutes a LDD event. Previous definitions of dispersal patterns emphasized only their distance components and characterized LDD events basically in terms of geographic distance between a dispersed propagule (or an established early seedling) and its most likely maternal or paternal (in case of pollen) source. Absolute and proportional definitions for the LDD events have been proposed depending on arbitrary thresholds of either the distance beyond which a dispersal event is LDD or the proportion of events occurring beyond a specific distance \citep{Nathan:2005jc, Nathan:2008is}. Thus, two key biological aspects of LDD events involve the transport of propagules outside a reference area: moving away from the source stand or population, and moving away from the area where relatives stand \citep{Kinlan:2005fb}. These two movements do not necessarily concur: a propagule may move over a very long distance yet still be disseminated within the reach of the neighborhood where parental individuals mate. Within a demo-genetic framework it is easy to envision a combination of situations concerning the spatial scale of the dispersal processes (Table 1) and unambiguously define different types of LDD events. The idea that dispersal occurs in reference to these two spatial reference frames, i.e., the population or stand and the genetic neighborhood area, is motivated by the fact that dispersal entails the movement of both an individual propagule (i.e., a pollen grain or a seed) and a distinct set of genes (i.e., the male genotype in case of pollen, or a seed genotype). Thus, dispersal entails simultaneous demographic and genetic effects through recruitment of new individuals in the population and through contributions to gene flow \citep{Harper:1977aa}. When considered its combined influence on demography and population genetics, the concept of LDD nicely bridges these two paradigms embedded in the biological definition of population \citep{Waples:2006ev}.

@@

While both $SDD_{loc}$ and $LDD_{loc}$ can be crucial for assuring the local persistence of populations, $LDD_{neigh}$ and $LDD_{ss}$ would be extremely important contributors to the structuring of genetic pools, realized gene flow, and maintaining connectivity in metapopulation scenarios. We argue that both the demographic and the genetic references are relevant for a proper definition of LDD.

@@

Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species. It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events \citep{Nathan:2008is,Schurr2009long}. Yet our current understanding of the extent, frequency, and consequences of LDD is very limited. On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD. On the other hand, we still have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.

@@

We advocate \citep[also see ][]{Jordano:2002tr,Nathan:2003qe,Jones:2008il,Hardesty:2011jn} a combination of approaches including large-scale biotelemetry to characterize animal movement, coupled with large-scale genetic sampling of dispersed propagules, and demogenetic approaches that combine both demographic and genetic research. A crucial aspect would be to effectively associate the role of individual frugivore species to specific dispersal outcomes, by identifying the actual disperser contributing a dissemination event \citep{GonzalezVaro:2014ij} and simultaneously characterizing the source maternal plant \citep{Jordano:2002tr}.

>>>>

1. Dispersal is a key individual-based process influencing many life-history attributes, scaling up to population-level properties (e.g., metapopulation connectivity). A persistent challenge in dispersal ecology has been the robust characterization of dispersal functions (kernels), a fundamental tool to predict how dispersal processes respond under global change scenarios. Especially the rightmost tail of these functions, i.e. the long-distance dispersal (LDD) events, are difficult to characterize empirically and to model in realistic ways. \\
2. But, when is it a LDD event? In the specific case of plants, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event. Here I discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration. \\
3. Strict-sense long distance dispersal involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals. Combinations of propagule movements within/outside these two spatial reference frames results in four distinct modes of LDD. \\
4. \textit{Synthesis.} I expect truncation of seed dispersal kernels to have multiple consequences on demography and genetics, following to the loss of key dispersal services in natural populations. Irrespective of neighborhood sizes, loss of LDD events may result in more structured and less cohesive genetic pools, with increased isolation-by-distance extending over broader areas. Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.\\

@@

I aim at providing a general framework for the quantitative analysis of LDD events so that estimates of their frequency and extent could be comparable across different study systems. I argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses. These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD. I start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds. I then use a case study to assess differential contributions of animal frugivores performing LDD.

Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events. I propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives. I also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds. Finally, I highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.

@@

Here I propose an explicit definition of LDD and what constitutes a LDD event. Previous definitions of dispersal patterns emphasized only their distance components and characterized LDD events basically in terms of geographic distance between a dispersed propagule (or an established early seedling) and its most likely maternal or paternal (in case of pollen) source. Absolute and proportional definitions for the LDD events have been proposed depending on arbitrary thresholds of either the distance beyond which a dispersal event is LDD or the proportion of events occurring beyond a specific distance \citep{Nathan:2005jc, Nathan:2008is}. Thus, two key biological aspects of LDD events involve the transport of propagules outside a reference area: moving away from the source stand or population, and moving away from the area where relatives stand \citep{Kinlan:2005fb}. These two movements do not necessarily concur: a propagule may move over a very long distance yet still be disseminated within the reach of the neighborhood where parental individuals mate. Within a demo-genetic framework it is easy to envision a combination of situations concerning the spatial scale of the dispersal processes (Table 1) and unambiguously define different types of LDD events. The idea that dispersal occurs in reference to these two spatial reference frames, i.e., the population or stand and the genetic neighborhood area, is motivated by the fact that dispersal entails the movement of both an individual propagule (i.e., a pollen grain or a seed) and a distinct set of genes (i.e., the male genotype in case of pollen, or a seed genotype). Thus, dispersal entails simultaneous demographic and genetic effects through recruitment of new individuals in the population and through contributions to gene flow \citep{Harper:1977aa}. When considered its combined influence on demography and population genetics, the concept of LDD nicely bridges these two paradigms embedded in the biological definition of population \citep{Waples:2006ev}.

@@

While both $SDD_{loc}$ and $LDD_{loc}$ can be crucial for assuring the local persistence of populations, $LDD_{neigh}$ and $LDD_{ss}$ would be extremely important contributors to the structuring of genetic pools, realized gene flow, and maintaining connectivity in metapopulation scenarios. I argue that both the demographic and the genetic references are relevant for a proper definition of LDD.

@@

Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species. It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events \citep{Nathan:2008is,Schurr2009long}. Yet our current understanding of the extent, frequency, and consequences of LDD is very limited. On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD. On the other hand, we have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.

@@

I advocate \citep[also see ][]{Jordano:2002tr,Nathan:2003qe,Jones:2008il,Hardesty:2011jn} a combination of approaches including large-scale biotelemetry to characterize animal movement, coupled with large-scale genetic sampling of dispersed propagules, and demogenetic approaches that combine both demographic and genetic research. A crucial aspect would be to effectively associate the role of individual frugivore species to specific dispersal outcomes, by identifying the actual disperser contributing a dissemination event \citep{GonzalezVaro:2014ij} and simultaneously characterizing the source maternal plant \citep{Jordano:2002tr}.
